Forecasting the Support of Kamala Harris in 2024 U.S. Presidential Election using a Model-Based Approach with Poll-of-Polls Data

### 2024-11-04 paper/ForecastingtheSupportofKamalaHarrisin2024U.S.PresidentialElection.qmd
update with wrong word usage

## Outcome variable {#sec-outcome-variable}

@@

```{r}
#| label: fig-percentage
#| fig-cap: Distribution of the Support Percentage of Kamala Harris
#| echo: false
#| warning: false
#| message: false
ggplot(harris_data, aes(x = pct)) +
  geom_histogram(binwidth = 1, fill = "gold", color = "grey50") +
  labs(
    x = "Support Percentage (%)",
    y = "Count"
  ) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

@@

@fig-numericgrade displays the distribution of numeric grades assigned
to various pollsters, with the x-axis representing the numeric grade
values and the y-axis indicates the frequency of each grade. The
numeric grade is a metric that evaluates the quality or reliability of a
pollster, taking into account factors such as methodology, historical
accuracy, sample size, and pollster reputation. In this chart, we
observe that the most common numeric grades are concentrated around 2.75
and 3.00, with a large spike at these values, suggesting that a
significant number of polls are conducted by highly rated pollsters.
Fewer polls have grades below 2.5, indicating a relatively lower
occurrence of polls from less-reliable pollsters in this dataset. This
distribution underscores the emphasis on high-quality pollsters within
the dataset, ensuring a more reliable and consistent source of polling
data for subsequent analysis.

ggplot(harris_data, aes(x = numeric_grade)) +
  geom_histogram(binwidth = 0.05, fill = "gold", color = "grey30") +
  labs(
    x = "Numeric Grade",
    y = "Frequency"
  ) +
  theme_minimal()
```

@@

ggplot(filtered_data, aes(x = end_date, y = pct, color = pollster)) +
  geom_point(alpha = 0.7) +
  geom_smooth(se = FALSE)+
  scale_color_viridis_d() +
  labs(y = "Harris Support Percent", x = "Date") +
  theme_classic() +
  theme(legend.position = "bottom", legend.title = element_blank())
```

@@

```{r}
#| label: fig-numericgradetime
#| fig-cap: Harris Support Over Time by Numeric Grade of the polls
#| echo: false
#| warning: false
#| message: false

ggplot(harris_data, aes(x = end_date, y = pct, color = factor(numeric_grade))) +
  geom_point(alpha = 0.7) +
  geom_smooth(se = FALSE, span = 0.3) +
  labs(y = "Support for Harris", x = "Date", color = "Numeric Grade") +
  scale_color_viridis_d() +
  theme_classic() +
  theme(legend.position = "bottom") +
  coord_cartesian(ylim = c(0, 100))
```

@@

The linear regression models constructed for predicting support for
Kamala Harris and Donald Trump in the 2024 U.S. Presidential Election
utilized both unweighted and weighted configurations. The model summary of support for Harris is shown in @tbl-summary1, and support for Trump is shown in @tbl-summary2. As shown in
@tbl-summary1 and @tbl-summary2, the intercepts for Harris and Trump,
estimated at 44.70% and 59.28% respectively in the unweighted models, while in weighted models, it's estimated at 45.2% and 54.8%,
represent their baseline support levels when all predictors are at
reference values. The high ( R\^2 ) values of 0.824 for Harris’s
unweighted model and 0.892 for her weighted model, along with a similar
values of 0.849 and 0.916 for Trump, indicate that a significant portion
of the variation in support can be explained by the model's predictors.
The relatively low Root Mean Square Error (RMSE) values further suggest
that the models provide reasonably accurate predictions of support
levels.

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| tbl-cap: "Summary of Multiple Linear Regression Model Predicting Harris Support Based on Polling Data"
#| label: tbl-summary1
msummary(
  list("Unweighted" = harris_unweighted_model, "Weighted" = harris_weighted_model),
  fmt = fmt_significant(3),coef_omit="state|pollster",notes = "Note. All models include a control for state and pollsters. The reference level of state is Alaska, the reference level of population is likely voter, and the reference level of pollsters is Alaska is Angus Reid.")
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| tbl-cap: "Summary of Multiple Linear Regression Model Predicting Trump Support Based on Polling Data"
#| label: tbl-summary2
msummary(
  list("Unweighted" = trump_unweighted_model, "Weighted" = trump_weighted_model),
  fmt = fmt_significant(3),coef_omit="state|pollster",notes = "Note. All models include a control for state and pollsters. The reference level of state is Alaska,the reference level of population is likely voter, and the reference level of pollsters is Alaska is ABC/Washington Post.")
```

>>>>

@fig-numericgrade displays the distribution of numeric grades assigned to various pollsters. The numeric grade is a metric that evaluates the quality or reliability of a pollster, considering factors such as methodology, historical accuracy, sample size, and pollster reputation. In this chart, we observe that the frequency of numeric grades peaks around 2.75 and 3.00, with a notable concentration at these values. This suggests that, after excluding lower-quality polls, a significant number of surveys are conducted by highly rated pollsters. Fewer polls have grades below 2.5, indicating a relatively lower occurrence of polls from less-reliable pollsters in this dataset. This distribution highlights the emphasis on high-quality pollsters, ensuring a more reliable and consistent source of polling data for subsequent analysis.

ggplot(harris_data, aes(x = numeric_grade)) +
  geom_histogram(binwidth = 0.05, fill = "gold", color = "grey30") +
  labs(
    x = "Numeric Grade",
    y = "Frequency"
  ) +
  theme_minimal()
```
## Outcome variable {#sec-outcome-variable}

@@

```{r}
#| label: fig-percentage
#| fig-cap: Distribution of the Support Percentage of Kamala Harris
#| echo: false
#| warning: false
#| message: false
ggplot(harris_data, aes(x = pct)) +
  geom_histogram(binwidth = 1, fill = "gold", color = "grey50") +
  labs(
    x = "Support for Harris (%)",
    y = "Count"
  ) +
  theme_classic() +
  theme(plot.title = element_text(hjust = 0.5))
```

@@

@@

ggplot(filtered_data, aes(x = end_date, y = pct, color = pollster)) +
  geom_point(alpha = 0.7) +
  geom_smooth(se = FALSE)+
  scale_color_viridis_d() +
  labs(y = "Support for Harris (%)", x = "End Date of the Poll") +
  theme_classic() +
  theme(legend.position = "bottom", legend.title = element_blank())
```

@@

```{r}
#| label: fig-numericgradetime
#| fig-cap: Harris Support Over Time by Numeric Grade of the Polls
#| echo: false
#| warning: false
#| message: false

ggplot(harris_data, aes(x = end_date, y = pct, color = factor(numeric_grade))) +
  geom_point(alpha = 0.7) +
  geom_smooth(se = FALSE, span = 0.3) +
  labs(y = "Support for Harris (%)", x = "End Date of the Poll", color = "Numeric Grade") +
  scale_color_viridis_d() +
  theme_classic() +
  theme(legend.position = "bottom") +
  coord_cartesian(ylim = c(0, 100))
```

@@

The linear regression models constructed for predicting support for
Kamala Harris and Donald Trump in the 2024 U.S. Presidential Election
utilized both unweighted and weighted configurations. The model summary of support for Harris is shown in @tbl-summary1, and support for Trump is shown in @tbl-summary2. As shown in
@tbl-summary1 and @tbl-summary2, the intercepts for Harris and Trump,
estimated at 44.70% and 59.28% respectively in the unweighted models, while in weighted models, it's estimated at 45.2% and 54.8%,
represent their baseline support levels when all predictors are at
reference values. The high ( $R^2$ ) values of 0.824 for Harris’s
unweighted model and 0.892 for her weighted model, along with a similar
values of 0.849 and 0.916 for Trump, indicate that a significant portion
of the variation in support can be explained by the model's predictors.
The relatively low Root Mean Square Error (RMSE) values further suggest
that the models provide reasonably accurate predictions of support
levels.

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| tbl-cap: "Summary of Multiple Linear Regression Model Predicting Harris Support Based on Polling Data"
#| label: tbl-summary1
msummary(
  list("Unweighted" = harris_unweighted_model, "Weighted" = harris_weighted_model),
  fmt = fmt_significant(3),coef_omit="state|pollster",notes = "Note. All models include a control for state and pollsters. The reference level of state is Alaska, the reference level of population is likely voter, and the reference level of pollsters is Angus Reid.")
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| tbl-cap: "Summary of Multiple Linear Regression Model Predicting Trump Support Based on Polling Data"
#| label: tbl-summary2
msummary(
  list("Unweighted" = trump_unweighted_model, "Weighted" = trump_weighted_model),
  fmt = fmt_significant(3),coef_omit="state|pollster",notes = "Note. All models include a control for state and pollsters. The reference level of state is Alaska,the reference level of population is likely voter, and the reference level of pollsters is ABC/Washington Post.")
```
